---
title: "class notes 1.28"
author: "Lucy Buhayenko"
format: html
embed-resources: true
---

Question: how can i hide output but still have the code run? How can I have output but not see the code?

Answer: use `results` to hide output but show code and `echo` to hide code but show output. Use `include` to execute code but not show it. Each of these can be true or false

```{r}
#| results: false
#| message: false

library(tidyverse)
library(readxl)

glimpse(economics)

```

As a general rule, include both code and output in Math 250 assignments.

So the above chunk isn't ideal from this perspective.

## Single-variable plots

We'll use the `erykah` data set again.

+ you need to complete the previous line -> you do esc and then go back to the previous line 

```{r}
erykah <- read_excel("erykah.xlsx")
```

The go-to plot for a single quantitative variable is the histogram.

Let's make one for `danceability`.

```{r}
ggplot(erykah, 
       aes(x = danceability)) + 
  geom_histogram() + 
  labs(title ="Danceability vs count") + theme_minimal() #for a cleaner look.
```


use built in color pallete and theme
