---
title: "class notes 2.9"
author: "Lucy Buhayenko"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
```



## A bit more on the filter. 

We saw how ot filter using "and" . Here's "or" using the `diamonds` data set:

```{r}
glimpse(diamonds)

diamonds_colorless <- filter(diamonds,
                              color == "D" | color == "E")

glimpse(diamonds_colorless)
```



When specifying multiple values of a single variabel, the `%in%` command can be helpful.

```{r}
filter(diamonds, color %in% c("D", "E", "F"))#will only keep the diamonds that are only D E or F
 c("D", "E", "F") # c makes a vector
```

## Subsetting columns with select

```{r}
select(diamonds,
       carat, 
       price,
       color) # you say what columns to keep - carat, price, color

select(diamonds,
       -carat) # specify columns to omit
```

Select is very flexible. 
```{r}
select(diamonds,
       table:z) #columns in a range

select(diamonds,
       2:6) #columns 2-6
```

In general, you should try to avoid specifying columns by number. It's less reliable (stable)

Some useful helper functions: 
```{r}
select(diamonds,
       starts_with("c")) #column name starts with c; this is good for 'caught' in scooby data set
```

Also, useful are `ends_with` and `contains`.

`Select` only works for columns. For rows use `filter`


