---
title: "Homework 3"
author: "Lucy Buhayenko"
format: html
embed-resources: true
code-fold: true
---

```{r}
#| label: setup
#| message: false


library(tidyverse)
library(modeldata)
```

## Problem 1
#### Recreated `Learn to Earn` box-plot
```{r}
#| label: problem 1
ggplot(attrition,
       aes(x = MonthlyIncome,
           y = Education,
           fill = Education)) +
  labs(x = "Monthly Income", 
       y = "Education",
       title = "Learn to Earn") +
  geom_boxplot(show.legend = FALSE) + 
  scale_fill_brewer(palette = "Dark2")
```



## Problem 2
#### Recreated age histogram plot
```{r}
#| label: problem 2
ggplot(attrition, 
       aes (x = Age)) + 
  geom_histogram(bins = 10,
                 boundary = 7,
                 color = "darkgreen",
                 fill = "lightgreen") + 
  theme_minimal()
```

## Problem 3
#### Bar chart stressing the different levels of job satisfaction but also showing the relative attrition rates within each group

```{r}
#| label: problem 3
ggplot(attrition, 
       aes(x = JobSatisfaction,
           fill = Attrition)) + 
  labs(x = "Job Satisfaction",
       y = "Frequency",
       title = "Levels of Job Satisfaction with relative attrition rates") + 
  geom_bar() + 
  scale_fill_brewer(palette = "Dark2") + 
  theme_minimal()
```

## Problem 4
#### Bar chart clearly comparing the different attrition rates for various levels of job satisfaction.

```{r}
#| label: problem 4
ggplot(attrition, 
       aes(x = JobSatisfaction,
       fill = Attrition)) + 
  labs(x = "Job Satisfaction",
       y = "Frequency",
       title = "Attrition Rates for Various Levels of Job Satisfaction") + 
  geom_bar(position = "fill") + 
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal()
```


## Problem 5
#### Recreated `Income by field` plot
```{r}
#| label: problem 5
ggplot(attrition,
       aes(x = MonthlyIncome,
           fill = EducationField)) +
  labs(x = "Monthly Income",
       y ="count", 
       title = "Income by field",
       caption = "Source: IBM Watson Analytics Lab") + 
  geom_histogram(binwidth = 655,
    show.legend = FALSE) + 
  facet_wrap(~EducationField,
             ncol = 2) + 
  scale_fill_brewer(palette = "Dark2") + 
  theme_minimal()
```



